 

### 2021-07-05 vertical-thermal-review.Rmd
Update vertical-thermal-review.Rmd

Greater $VPD$, atmospheric demand and low RH can be more stressful for trees, than soil moisture across biomes-including mesic forests,  as the former limits surface conductance and evapotranspiration to a greater extent than soil moisture [@novickIncreasingImportanceAtmospheric2016; @corlettImpactsWarmingTropical2011; @ruehrWaterAvailabilityDominant2016]. Along with greater VPD, increase in temperature even under adequete water can cause greater tree hydraulic stress due to altered soil-water dynamics than milder drought, in temperate species [@ruehrWaterAvailabilityDominant2016; @parkwilliamsTemperaturePotentDriver2013]. Whereas greater $RH$ can do the opposite, stifling evapotranspiration and thereby increasing $T_{leaf}$ [Fig.3; @perezIncreasingHumidityThreatens2018; @tibbittsHumidityPlants1979]. Altered precipitation regimes can cause droughts, that are more  detrimental to larger trees than small trees, because limited soil-water cannot meet the evaporative demand via the long-distance water transport to the canopy [@bennettLargerTreesSuffer2015; @mcgregor_tree_nodate; @bartholomewSmallTropicalForest2020]. However, given adequate moisture to sustain high $g_s$, canopy leaves exposed to more wind and lower humidity should be more effective at cooling when $T_{air}$ exceeds physiological optima (Fig. `r fig_leaf_T`).

>>>>

Greater $VPD$ and low RH can be more stressful for trees, than soil moisture across biomes-including mesic forests, as the former limits surface conductance and evapotranspiration to a greater extent than soil moisture [@novickIncreasingImportanceAtmospheric2016; @corlettImpactsWarmingTropical2011; @ruehrWaterAvailabilityDominant2016]. Whereas greater $RH$ can do the opposite, stifling evapotranspiration and thereby increasing $T_{leaf}$ [Fig.3; @perezIncreasingHumidityThreatens2018; @tibbittsHumidityPlants1979]. Altered precipitation regimes can cause droughts, that are more  detrimental to larger trees than small trees, because limited soil-water cannot meet the evaporative demand via the long-distance water transport to the canopy [@bennettLargerTreesSuffer2015; @mcgregor_tree_nodate; @bartholomewSmallTropicalForest2020]. However, given adequate moisture to sustain high $g_s$, canopy leaves exposed to more wind and lower humidity should be more effective at cooling when $T_{air}$ exceeds physiological optima (Fig. `r fig_leaf_T`).
